\chapter{Introduction}
Path planning, collision detection, localization, mapping, controlling of systems and many others are challenging problems in robotics.
Moreover, these problems have to be often solved in real-time applications which makes them even more challenging.
To be able to solve them efficiently, these problems are typically formulated as systems of polynomial equations and inequalities or as polynomial optimization tasks.

The state of the art methods for solving polynomial systems are based on polynomial algebra, i.e.\ on Gr\"obner bases computation and computation of eigenvalues and eigenvectors of multiplication matrices \cite{Cox-Little-Shea2015}.
Contrary to this approach, in this work, we will use methods of non-linear optimization, especially techniques from semidefinite programming \cite{Lasserre}, to solve the polynomial systems.
This is a novel idea in the field of robotics.
Moreover, we can take advantage of these methods, because we are no longer limited to systems of polynomial equations, but we can also take into account polynomial inequalities.
These methods also allow us to solve polynomial optimization problems, i.e.\ to find a global optimal value of a polynomial function with given polynomial constraints.

Moreover, some non-real solutions often appear as a side effect amongst real solutions when a geometric task is modeled as a polynomial problem.
These superfluous non-real solutions are not solutions of the original geometric problem and are not relevant for us.
Unfortunately, the state of the art methods from algebraic geometry can not be limited to find real solutions only.
The contemporary approach is to find all complex solutions of the polynomial problem and filter the non-real solutions out afterward.
With the use of the techniques from semidefinite programming \cite{Laurent2012} we can limit ourselves to real numbers only, and therefore save some computation time by not computing the non-real solutions.
