\section{Polynomial algebra preliminaries}
To make this chapter self-contained, we introduce the minimal possible amount of notation for polynomial algebra.
The overall review of the key concepts can be seen in \cite{Cox-Little-Shea2015}, which notation we will follow.

\subsection{The polynomial ring, ideals and varieties}
In this section, we will consider polynomials in $n$ unknowns $x_1, x_2, \ldots, x_n$ with real coefficients.

\begin{definition}[Monomial {\thecite[Chapter~1, §1, Definition~1]{Cox-Little-Shea2015}}]
  A monomial in variables $x_1, x_2, \ldots , x_n$ is a product of the form
  \begin{align}
    x_1^{\alpha_1} x_2^{\alpha_2} \cdots x_n^{\alpha_n},
  \end{align}
  where all of the exponents $\alpha_1, \alpha_2, \ldots, \alpha_n \in \Zpn$.
  The total degree of this monomial is the sum $\alpha_1 + \alpha_2 + \cdots + \alpha_n$.
\end{definition}

When we let $\alpha = \begin{pmatrix}\alpha_1 & \alpha_2 & \cdots & \alpha_n\end{pmatrix}$ to be an $N$-tuple then we can simplify the notation for monomials by setting
\begin{align}
  x^\alpha &= x_1^{\alpha_1} x_2^{\alpha_2} \cdots x_n^{\alpha_n}.
\end{align}
We call the $N$-tuple $\alpha$ the multidegree of the monomial $x^\alpha$.
Let us also denote $|\alpha| = \alpha_1 + \alpha_2 + \cdots + \alpha_n$ the total degree of the monomial $x^\alpha$.
Note that $x^\alpha = 1$ when $\alpha = \begin{pmatrix}0 & 0 & \cdots & 0\end{pmatrix}$.

\begin{definition}[Polynomial {\thecite[Chapter~1, §1, Definition~2]{Cox-Little-Shea2015}}]
  A polynomial $f$ in variables $x_1, x_2, \cdots, x_n$ with real coefficients is a finite linear combination (with real coefficients) of monomials.
  We will write a polynomial $f$ in the form
  \begin{align}
    f &= \sum_{\alpha\in\Zpn^n} c_\alpha x^\alpha,
  \end{align}
where $c_\alpha\in\R$, and there are only finitely many terms $c_\alpha x^\alpha$ in the sum.
\end{definition}

We call $c_\alpha$ the coefficient of the monomial $x^\alpha$.
If $c_\alpha \neq 0$, then we call $c_\alpha x^\alpha$ a term of the polynomial $f$.
The total degree of the polynomial $f$ is denoted as $\deg(f) = \max_{\alpha\in\Zpn^n} |\alpha|$ for non-zero coefficients $c_\alpha\in\R$.
The zero polynomial has no degree.
The set of all polynomials in variables $x = \begin{bmatrix}x_1 & x_2 & \cdots & x_n\end{bmatrix}^\top$ with real coefficients is the ring of multivariate polynomials denoted as $\R[x]$.

\begin{example}
  Let us have polynomial $f \in\R[x, y, z]$ in three variables:
  \begin{align}
    f &= x^3y - 4xz^2 + \frac{2}{3}z^2 + 1.
  \end{align}
  The polynomial $f$ is a sum of four terms.
  The second term $-4xz^2$ with multidegree $\alpha = \begin{pmatrix}1 & 0 & 2\end{pmatrix}$ has a coefficient $-4$ and monomial $xz^2$.
  The total degree of this monomial is $1 + 0 + 2 = 3$, but the total degree $\deg(f)$ of the polynomial $f$ is 4, because that is the multidegree of the monomial $x^3y$.
\end{example}

\begin{definition}[Ideal {\thecite[Chapter~1, §4, Definition~1]{Cox-Little-Shea2015}}]
  A subset $\Ideal \subseteq \Rx$ is an ideal if it satisfies:
  \begin{enumerate}
    \item $0 \in \Ideal$.
    \item If $f, g \in \Ideal$, then $f + g \in \Ideal$.
    \item If $f \in \Ideal$ and $h \in \Rx$, then $hf \in \Ideal$.
  \end{enumerate}
\end{definition}

\begin{definition}[{\thecite[Chapter~1, §4, Definition~2]{Cox-Little-Shea2015}}]
  Let $f_1, f_2, \ldots, f_m$ be polynomials in $\Rx$.
  Then we set
  \begin{align}
     \langle f_1, f_2, \ldots, f_m\rangle &= \Bigg\{\sum_{j=1}^mh_jf_j\ |\ h_1, h_2, \ldots, h_m\in\Rx\Bigg\}.
  \end{align}
\end{definition}

The set $\langle f_1, f_2, \ldots, f_m\rangle$ is an ideal and we call it the ideal generated by $f_1, f_2, \ldots, f_m$.

\begin{definition}[Radical ideal {\thecite[Chapter~4, §2, Definition~2]{Cox-Little-Shea2015}}]\labeldef{pol:prelim:radical}
  An ideal $\Ideal$ is radical if $f^m \in \Ideal$ for some integer $m \in\Zp$ implies that $f \in \Ideal$.
\end{definition}

\begin{definition}[Radical of ideal {\thecite[Chapter~4, §2, Definition~4]{Cox-Little-Shea2015}}]
  Let $\Ideal \subseteq \Rx$ be an ideal.
  The radical of the ideal \Ideal, denoted \rIdeal, is the set
  \begin{align}
    \rIdeal &= \big\{f\in\R[x]\ |\ f^m\in \Ideal \text{ for some } m\in\Zp\big\}.
  \end{align}
\end{definition}

%TODO Real Radical Ideal

Note that it always holds that $\Ideal \subseteq \rIdeal$.
The reason is that taking every polynomial $f \in \Ideal$, \ie{} $f^1 \in \Ideal$, which implies by definition that $f$ is in \rIdeal.
Moreover, $\Ideal = \rIdeal$ if and only if the ideal $\Ideal$ is radical.

\begin{definition}[Affine variety {\thecite[Chapter~1, §2, Definition~1]{Cox-Little-Shea2015}}]
  Let $f_1, f_2, \ldots, f_m$ be polynomials in \Rx.
  Then we set
  \begin{align}
    \VarC(f_1, f_2, \ldots, f_m) &= \big\{\xi \in \C^n\ |\ f_i(\xi) = 0\text{ for all }1 \leq i \leq m\big\}.
  \end{align}
  We call $\VarC(f_1, f_2, \ldots, f_m)$ the affine variety defined by the polynomials $f_1, f_2, \ldots, f_m$.
\end{definition}

%TODO Real affine variety

\begin{definition}[Vanishing ideal of affine variety {\thecite[Chapter~1, §4, Definition~5]{Cox-Little-Shea2015}}]
  Let $\Var \subseteq \C^n$ be an affine variety. Then we set the vanishing ideal $\Ideal(\Var)$ of affine variety $\Var$ as
  \begin{align}
    \Ideal(\Var) &= \{f \in \Rx \ |\ f(\xi) = 0 \text{ for all } \xi \in \Var\}.
  \end{align}

\end{definition}

%TODO Nullstellensatzen
%pol:prelim:nss
%pol:prelim:realnss

Let us better explain the defined objects and their relations in the following example.

\begin{example}
  Consider the following polynomial $f\in\Rx$
  \begin{align}
    f(x) = (x-1)^2(x+3)(x^2+1)
  \end{align}
  and the ideal $\Ideal = \langle f \rangle$ generated by the polynomial.
  The ideal \Ideal{} contains the polynomial $f$, but also all polynomial multiples of $f$, say in the form $g_j = h_jf$, eg.\ for $h_1 = x^3$, $g_1(x) = x^3(x-1)^2(x+3)(x^2+1)$ or for $h_2 = (x+3)(x^2+1)$, $g_2(x) = (x-1)^2(x+3)^2(x^2+1)^2$.

  The ideal \Ideal{} is not radical, because e.g.\ polynomial $g_2$ is in \Ideal{}, but polynomial $(x-1)(x+3)(x^2+1)$ is not and that violates the condition given in \refdef{pol:prelim:radical}.

  The affine variety \VarC{} of the ideal \Ideal{} are all points $\xi \in \C^n$ on which all the polynomials from \Ideal{} evaluates to zero. In our case, these points are the roots of the polynomial $f$. Therefore,
  \begin{align}
    \VarC(\Ideal) = \{1, -3, -i, i\},
  \end{align}
  where the root 1 has multiplicity two and all the other roots have multiplicity one.
  We can easily see, that all polynomials $h_jf$ from \Ideal{} vanishes on $\xi \in \VarC$, since $h_j(\xi)f(\xi) = h_j(\xi)0 = 0$ as $\xi$ is a root of $f$.

  Using the \refthe{pol:prelim:nss} we can now construct the radical of \Ideal{}.
  The radical \rIdeal{} contains all polynomials that vanish on the points $\VarC(\Ideal)$, i.e. all polynomial multiples of a polynomial with roots $\VarC$, taken all with multiplicity one.
  We can therefore write that the radical is generated by a single polynomial $f_{\sqrt{}}(x) = (x-1)(x+3)(x^2+1)$.
  \begin{align}
    \rIdeal &= \langle (x-1)(x+3)(x^2+1) \rangle
  \end{align}
  Now, \rIdeal{} is satisfying the condition in \refdef{pol:prelim:radical} for a radical.
  Note, that $\Ideal{} \subset \rIdeal{}$.

  Let us make the same exercise for polynomials with real roots.
  The real affine variety \VarR(\Ideal) is the intersection of the affine variety $\VarC(\Ideal)$ with real numbers, i.e.\
  \begin{align}
    \VarR(\Ideal) &= \{1, -3\}.
  \end{align}

  Using the \refthe{pol:prelim:realnss} we can now construct the real radical of \Ideal{}.
  As before, we construct the real radical $\rrIdeal{}$ as an ideal generated by one polynomial, which has only roots from $\VarR$ all with multiplicity one.
  Such a polynomial is $f_{\sqrt[\R]{}}(x) = (x-1)(x+3)$.
  \begin{align}
    \rrIdeal &= \langle(x-1)(x+3) \rangle
  \end{align}
  Now, let us check that \rrIdeal{} is really real radical of \Ideal.
  We show, that \refeq{pol:prelim:???} holds for atleast one polynomial from \rrIdeal.
  Let us see, for example, if $f_{\sqrt[\R]{}}^{2m} + h^2 \in \Ideal$.
  This hold for $m = 1$ and $h(x) = x(x-1)(x+3)$.
  We are getting:
  \begin{align}
    f_{\sqrt[\R]{}}^{2m}(x) + h^2(x) &= (x-1)^2(x+3)^2 + x^2(x-1)^2(x+3)^2\\
    &= (x-1)^2(x+3)^2(x^2+1)\\
    &= (x+3)f(x)
  \end{align}
  Where $(x+3)f(x)$ is in \Ideal{} as it is a polynomial multiple of $f(x)$.

\end{example}
